---
title: "4th homework"
author: "Manvydas Sokolovas, p3170190"
date: "11/30/2017"
output:
  pdf_document:
    latex_engine: xelatex
    number_section: yes
    toc_depth: 2
  html_document: default
---

```{r, include=FALSE}
library(car)
library(knitr)
library(boot)

balt <- read.csv("balt.csv"); balt <- balt[,-1]
```

# 1st regression summary
```{r, echo = FALSE}
model1 <- lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC, data = balt)
kable(summary(model1)$coef)
kable(data.frame(summary(model1)[8]))
```

## Bootstrap for the first regression
```{r, echo = FALSE}
bootfun <- function(data, index) {
  coef(lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC, data = balt, subset = index))
}

b1<-boot(balt, bootfun, 10000)
b1
```

## Bootstrap for $R^2$ confidence intervals
```{r, echo = FALSE}
R2 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC, data[indices,]))$r.squared, R = 10000)

R2
quantile(R2$t,c(0.025,0.975))
```

# Confidence intervals of the first regression estimates
```{r, echo = FALSE}
conf<-data.frame()
for(i in 1:6){
  for(l in 1:2){
   conf[i,l] <- round(quantile(b1$t[,i],c(0.025,0.975))[l],3)
  }
}
colnames(conf) <- c("2.5%", "97.5%")
rownames(conf) <- names(model1$coefficients)
kable(conf)
```

# 2nd regression summary
```{r, echo = FALSE}
mod2 <- lm(PRICE ~ NROOM + NBATH + PATIO + AC, data = balt)
kable(summary(mod2)$coef)
kable(data.frame(summary(mod2)[8]))
```

## Bootstrap for the second regression
```{r, echo = FALSE}
bootfun2 <- function(data, index) {
  coef(lm(PRICE ~ NROOM + NBATH + PATIO + AC, data = balt, subset = index))
}

b2<-boot(balt, bootfun2, 10000)
b2
```

## Bootstrap to get $R^2$ confidence interval
```{r, echo = FALSE}
R22 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + AC, data[indices,]))$r.squared, R = 10000)

R22
quantile(R22$t,c(0.025,0.975))
```

##
```{r, echo = FALSE}
box <- data.frame(R2$t, R22$t)
colnames(box) = c("model1","model2")
boxplot(box)  
a <- kable(data.frame(summary(model1)[8])) # R^2 firts model
b <- kable(data.frame(summary(mod2)[8])) # R^2 second model
```

* In the second model ${R}^2$ is smaller

* ${R}^2$ confidence intervals from both models are overlapping, but not a lot.

* Standart errors in the second model got bigger. Thus, assuming all those statements I would prefer the first model.

## Confidence intervals of second model estimates
```{r, echo = FALSE}
conf2<-data.frame()
for(i in 1:5){
  for(l in 1:2){
   conf2[i,l] <- round(quantile(b2$t[,i],c(0.025,0.975))[l],3)
  }
}
colnames(conf2) <- c("2.5%", "97.5%")
rownames(conf2) <- names(mod2$coefficients)
kable(conf2)
```

# Better models by R-square

```{r, message = FALSE, echo = FALSE}
panel.hist <- function(x, ...)    #ši funkcija reikalinga grafikų lentelei išbrėžti (histogramos pateikimui)                        
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor2 <- function(x, y, digits=2, cex.cor)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  test <- cor.test(x,y)
  Signif <- ifelse(round(test$p.value,3)<0.001,"p<0.001",paste("p=",round(test$p.value,3)))  
  text(0.5, 0.25, paste("r=",txt))
  text(.5, .75, Signif)
}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor,...)    #ši funkcija reikalinga grafikų lentelei
  #išbrėžti (koreliacijos koeficiento radimui)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = 3)
}
```

```{r, include = FALSE, echo = FALSE}
pairs(balt,upper.panel=panel.smooth,diag.panel=panel.hist, lower.panel=panel.cor2)
print(cor(balt), digits = 2)
```

*mod1* = ${PRICE} = {\beta}_0 + {NROOM} {\beta}_1 + {NBATH} {\beta}_2 + {PATIO} {\beta}_3 + {FIREPL} {\beta}_4 + {AC} {\beta}_5$

*mod2* = ${PRICE} = {\beta}_0 + {NROOM} {\beta}_1 + {NBATH} {\beta}_2 + {PATIO} {\beta}_3 + {AC} {\beta}_4$

*mod3* =  ${PRICE} = {\beta}_0 + {NBATH} {\beta}_1 + {PATIO} {\beta}_2 + {FIREPL} {\beta}_3 + {SQFT} {\beta}_4$

*mod4* =  ${PRICE} = {\beta}_0 + {PATIO} {\beta}_1 + {FIREPL} {\beta}_2 + {AC} {\beta}_3 + {BMENT} {\beta}_4 + {GAR} {\beta}_5 + {AGE} {\beta}_6 + {SQFT} {\beta}_7$

*mod5* =  ${PRICE} = {\beta}_0 + {NROOM} {\beta}_1 + {NBATH} {\beta}_2 + {PATIO} {\beta}_3 + {FIREPL} {\beta}_4 + {AC} {\beta}_5 + {GAR} {\beta}_6 + {AGE} {\beta}_7 + {SQFT} {\beta}_8$

*mod6* =  ${PRICE} = {\beta}_0 + {NBATH} {\beta}_1 + {PATIO} {\beta}_2 + {FIREPL} {\beta}_3 + {AC} {\beta}_4 + {GAR} {\beta}_5 + {AGE} {\beta}_6 + {SQFT} {\beta}_7$

*mod_all* =  ${PRICE} = {\beta}_0 + {NROOM} {\beta}_1 + {NBATH} {\beta}_2 + {PATIO} {\beta}_3 + {FIREPL} {\beta}_4 + {AC} {\beta}_5 + {BMENT} {\beta}_6 + {NSTOR} {\beta}_7 + {GAR} {\beta}_8 + {AGE} {\beta}_9 + {SQFT} {\beta}_10$


NROOM + NBATH + PATIO + FIREPL + AC + BMENT + NSTOR + GAR + AGE + SQFT

```{r, echo = FALSE}
mod1 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC, data[indices,]))$r.squared, R = 3000)

mod2 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + AC, data[indices,]))$r.squared, R = 3000)

mod3 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NBATH + PATIO + FIREPL + SQFT, data[indices,]))$r.squared, R = 3000)

mod4 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ PATIO + FIREPL + AC + BMENT + GAR + AGE + SQFT, data[indices,]))$r.squared, R = 3000)

mod5 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC + GAR + AGE + SQFT, data[indices,]))$r.squared, R = 3000)

mod6 <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NBATH + PATIO + FIREPL + AC + GAR + AGE + SQFT, data[indices,]))$r.squared, R = 3000)

mod_all <- boot(balt,function(data,indices)
        summary(lm(PRICE ~ NROOM + NBATH + PATIO + FIREPL + AC + BMENT + NSTOR + GAR + AGE + SQFT, data[indices,]))$r.squared, R = 3000)

rr <- data.frame(mod1$t, mod2$t, mod3$t, mod4$t, mod5$t, mod6$t, mod_all$t)
colnames(rr) <- c("mod1", "mod2", "mod3", "mod4", "mod5", "mod6", "mod_all")

boxplot(rr)
```

* $mod5$ ${R}^2$ confidence interval is just a bit higher than $mod6$. But $mod5$ have one more variable. Thus, that could happen because ${R}^2$ gets bigger when you add variables.
